---
title: "Density and number significant nursery species"
author: 'Dirkje Verhoeven, Dolf Andringa, Annelies Andringa: Marine Conservation Philippines'
date: "03 June 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load('data.RDa') #load the data
library('vegan')
library('ggplot2')
library('PMCMR')
library('lmtest')
nursery_species <- subset(df.species_all, df.species_all$nursery) 
attach(df.sample_data)
```

## Calculate some info per sample

```{r calculation}
#calculate the Shannon-Wiener diversity (with diversity function from vegan package) for all species
df.sample_data$diversity <- apply(df.sample_data[df.species_all$colname], MARGIN=1, diversity)
#calculate the Shannon-Wiener diversity (with diversity function from vegan package) for nursery species
df.sample_data$nursery_diversity <- apply(df.sample_data[nursery_species$colname], MARGIN=1, diversity)

#calculate number of species for all species by summing the presence/absence columns for those species
df.sample_data$num_species <- apply(df.sample_data[c(df.species_all$colname_pres)],MARGIN=1,sum)
#calculate number of species for nursery species by summing the presence/absence columns for those species
df.sample_data$num_nursery_species <- apply(df.sample_data[c(nursery_species$colname_pres)],MARGIN=1,sum)

#calculate number of fish
df.sample_data$num_fish <- apply(df.sample_data[c(df.species_all$colname)],MARGIN=1,sum)
#calculate number of nursery fish
df.sample_data$num_nursery_fish <- apply(df.sample_data[c(nursery_species$colname)],MARGIN=1,sum)

df.sample_data <- merge(x=df.sample_data, y=df.sites, by.x="location", by.y="name", all.x=TRUE)
```

```{r subset_nursery_species}
nursery_columns <- c(c("location","depth","transect","date","nursery_diversity","num_nursery_species","num_nursery_fish"),nursery_species$colname, nursery_species$colname_pres) #columns for nursery species analysis
df.nursery_sample_data <- df.sample_data[nursery_columns] # subset of the data with only the nursery species
```

## Calculate diversity and number of nursery fish species (mangrove AND seagrass)

```{r diversity_sign_species}

sig_species <-df.species_all[c(2,25,28,36,39,40,42,47,52,53,54,58,66,88,94,95,96,97,98,99,102,103,104),]

#calculate the Shannon-Wiener diversity (with diversity function from vegan package) for SIGNIFICANT nursery species
df.sample_data$sig_nurs_diversity <- apply(df.sample_data[sig_species$colname], MARGIN=1, diversity)

#calculate the number of nursery fish species
df.sample_data$sig_nurs_species <- apply(df.sample_data[c(sig_species$colname_pres)],MARGIN=1,sum)

#calculate number of SIGNFICANT nursery fish
df.sample_data$sig_nurs_num <- apply(df.sample_data[sig_species$colname], MARGIN=1, sum)
```

```{r summary_per_site}
df.summarydata <- aggregate(df.sample_data[c('sig_nurs_num','sig_nurs_diversity','sig_nurs_species')], by=list(df.sample_data$location), FUN=function(x) c(mean=mean(x), sum=sum(x),sd=sd(x),n=length(x)))
df.summarydata <- do.call(data.frame, df.summarydata)
```
This is the result per significant nursery species
```{r display_summary_per_site, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}
knitr::kable(df.summarydata)
```

##GLM on diversity of nursery fish species (log(size) & log(distance))

Gaussian GLM, since sig_nurs_diversity is a diversity-value not abundance

```{r glm_div_loop}
  res <- glm(formula = sig_nurs_diversity ~ log(distance_seagrass) * log(size_seagrass) * log(distance_mangroves) * log(size_mangroves) * mpa * depth * north, data = df.sample_data)

  summary(res)

  par(mfrow=c(1,2)) 
  plot(fitted(res),residuals(res),xlab="Fitted values", ylab="Residuals", abline(h=0, lty=2))
  tryCatch(
    lines(smooth.spline(fitted(res), residuals(res))), 
    error=function(e){}  )
  qqnorm(res$residuals)
  qqline(res$residuals) 

```

``` {r likelihood_diversity}
  model1 <- glm(formula = sig_nurs_diversity ~ log(distance_seagrass) * log(size_seagrass) * log(distance_mangroves) * log(size_mangroves) * mpa * depth * north, data = df.sample_data)
  model2 <- glm(formula = sig_nurs_diversity ~ log(distance_seagrass) * log(size_seagrass) * log(distance_mangroves) * log(size_mangroves) * mpa * north, data = df.sample_data)
  lrtest(model1, model2)
  #>0.05 = model 2 predicts data significantly better than model 1 {eliminate variabele with highest p-value}
  #<0.05 = model 2 predicts data significantly less than model 1 {end result}
```

##GLM on number of nursery fish species

```{r glm_sp_loop}
  res <- glm(formula = sig_nurs_species ~ log(distance_seagrass) * log(size_seagrass) * log(distance_mangroves) * log(size_mangroves) * mpa * depth * north, data= df.sample_data)
  
  summary(res)
  
  par(mfrow=c(1,2)) 
  plot(fitted(res),residuals(res),xlab="Fitted values", ylab="Residuals", abline(h=0, lty=2))
  tryCatch(
    lines(smooth.spline(fitted(res), residuals(res))), 
    error=function(e){}  )
  qqnorm(res$residuals)
  qqline(res$residuals) 

```

``` {r likelihood_species}
  model1 <- glm(formula = sig_nurs_species ~ log(distance_seagrass) * log(size_seagrass) * log(distance_mangroves) * log(size_mangroves) * mpa * depth * north, data = df.sample_data)
  model2 <- glm(formula = sig_nurs_species ~ log(distance_seagrass) * log(size_seagrass) * log(distance_mangroves) * log(size_mangroves) * mpa * north, data = df.sample_data)
  lrtest(model1, model2)
  #>0.05 = model 2 predicts data significantly better than model 1 {eliminate variabele with highest p-value}
  #<0.05 = model 2 predicts data significantly less than model 1 {end result}
```

## Calculate diversity and number of nursery fish species (mangrove AND seagrass); without species only significant with interaction (size/distance*depth)

```{r diversity_sign_species2}

sig_species <-df.species_all[c(2,25,28,36,39,42,47,52,53,54,58,88,94,95,96,97,98,99,102,103,104),]

#calculate the Shannon-Wiener diversity (with diversity function from vegan package) for SIGNIFICANT nursery species
df.sample_data$sig_nurs_diversity <- apply(df.sample_data[sig_species$colname], MARGIN=1, diversity)

#calculate the number of nursery fish species
df.sample_data$sig_nurs_species <- apply(df.sample_data[c(sig_species$colname_pres)],MARGIN=1,sum)

#calculate number of SIGNFICANT nursery fish
df.sample_data$sig_nurs_num <- apply(df.sample_data[sig_species$colname], MARGIN=1, sum)
```

```{r summary_per_site2}
df.summarydata <- aggregate(df.sample_data[c('sig_nurs_num','sig_nurs_diversity','sig_nurs_species')], by=list(df.sample_data$location), FUN=function(x) c(mean=mean(x), sum=sum(x),sd=sd(x),n=length(x)))
df.summarydata <- do.call(data.frame, df.summarydata)
```
This is the result per significant nursery species
```{r display_summary_per_site2, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}
knitr::kable(df.summarydata)
```

##GLM on diversity of nursery fish species (log(size) & log(distance))

Gaussian GLM, since sig_nurs_diversity is a diversity-value not abundance

```{r glm_div_loop2}
  res <- glm(formula = sig_nurs_diversity ~ log(distance_seagrass) * log(size_seagrass) * log(distance_mangroves) * log(size_mangroves) * mpa * depth * north, data = df.sample_data)

  summary(res)

  par(mfrow=c(1,2)) 
  plot(fitted(res),residuals(res),xlab="Fitted values", ylab="Residuals", abline(h=0, lty=2))
  tryCatch(
    lines(smooth.spline(fitted(res), residuals(res))), 
    error=function(e){}  )
  qqnorm(res$residuals)
  qqline(res$residuals) 

```

``` {r likelihood_diversity2}
  model1 <- glm(formula = sig_nurs_diversity ~ log(distance_seagrass) * log(size_seagrass) * log(distance_mangroves) * log(size_mangroves) * mpa * depth * north, data = df.sample_data)
  model2 <- glm(formula = sig_nurs_diversity ~ log(distance_seagrass) * log(size_seagrass) * log(distance_mangroves) * log(size_mangroves) * mpa * north, data = df.sample_data)
  lrtest(model1, model2)
  #>0.05 = model 2 predicts data significantly better than model 1 {eliminate variabele with highest p-value}
  #<0.05 = model 2 predicts data significantly less than model 1 {end result}
```

##GLM on number of nursery fish species

```{r glm_sp_loop2}
  res <- glm(formula = sig_nurs_species ~ log(distance_seagrass) * log(size_seagrass) * log(distance_mangroves) * log(size_mangroves) * mpa * depth * north, data= df.sample_data)
  
  summary(res)
  
  par(mfrow=c(1,2)) 
  plot(fitted(res),residuals(res),xlab="Fitted values", ylab="Residuals", abline(h=0, lty=2))
  tryCatch(
    lines(smooth.spline(fitted(res), residuals(res))), 
    error=function(e){}  )
  qqnorm(res$residuals)
  qqline(res$residuals) 

```

``` {r likelihood_species_2}
  model1 <- glm(formula = sig_nurs_species ~ log(distance_seagrass) * log(size_seagrass) * log(distance_mangroves) * log(size_mangroves) * mpa * depth * north, data = df.sample_data)
  model2 <- glm(formula = sig_nurs_species ~ log(distance_seagrass) * log(size_seagrass) * log(distance_mangroves) * log(size_mangroves) * mpa * north, data = df.sample_data)
  lrtest(model1, model2)
  #>0.05 = model 2 predicts data significantly better than model 1 {eliminate variabele with highest p-value}
  #<0.05 = model 2 predicts data significantly less than model 1 {end result}
```

