---
title: "GLM on abundance_pres"
author: 'Dirkje Verhoeven, Dolf Andringa, Annelies Andringa: Marine Conservation Philippines'
date: "02 June 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load('data.RDa') #load the data
library('vegan')
library('ggplot2')
library('PMCMR')
nursery_species <- subset(df.species_all, df.species_all$nursery) 
```

## Calculate some info per sample

```{r calculation}
#calculate the Shannon-Wiener diversity (with diversity function from vegan package) for all species
df.sample_data$diversity <- apply(df.sample_data[df.species_all$colname], MARGIN=1, diversity)
#calculate the Shannon-Wiener diversity (with diversity function from vegan package) for nursery species
df.sample_data$nursery_diversity <- apply(df.sample_data[nursery_species$colname], MARGIN=1, diversity)

#calculate number of species for all species by summing the presence/absence columns for those species
df.sample_data$num_species <- apply(df.sample_data[c(df.species_all$colname_pres)],MARGIN=1,sum)
#calculate number of species for nursery species by summing the presence/absence columns for those species
df.sample_data$num_nursery_species <- apply(df.sample_data[c(nursery_species$colname_pres)],MARGIN=1,sum)

#calculate number of fish
df.sample_data$num_fish <- apply(df.sample_data[c(df.species_all$colname)],MARGIN=1,sum)
#calculate number of nursery fish
df.sample_data$num_nursery_fish <- apply(df.sample_data[c(nursery_species$colname)],MARGIN=1,sum)

df.sample_data <- merge(x=df.sample_data, y=df.sites, by.x="location", by.y="name", all.x=TRUE)
```

## Nursery species

```{r subset_nursery_species}
nursery_columns <- c(c("location","depth","transect","date","nursery_diversity","num_nursery_species","num_nursery_fish"),nursery_species$colname, nursery_species$colname_pres) #columns for nursery species analysis
df.nursery_sample_data <- df.sample_data[nursery_columns] # subset of the data with only the nursery species
```

##GLM on abundance_pres

```{r glm_nur_ab_pres_loop}

speciesrows <- c(1:3,25:28,36,39,40,42,47,52:54,58,66,84:86,88,91:100,102:104)
for (spnum in speciesrows){
  sp <- df.species_all[spnum,]
  print(sp$colname)
  n <- subset(df.sample_data, df.sample_data[c(sp$colname_pres)] > 0)
  if(nrow(n)<5){
    print("Less than 5 rows remaining in subset")
    next
  }
  if(length(unique(n$depth))==1 & length(unique(n$mpa))>1){
    formula <- n[,c(sp$colname)] ~ n$distance_seagrass * n$size_seagrass * n$mpa * n$north
  }
  else if (length(unique(n$depth))>1 & length(unique(n$mpa))==1){
    formula <- n[,c(sp$colname)] ~ n$distance_seagrass * n$size_seagrass * n$depth * n$north  
  }
  else if (length(unique(n$depth))==1 & length(unique(n$mpa))==1){
    formula <- n[,c(sp$colname)] ~ n$distance_seagrass * n$size_seagrass * n$north  
  }
  else{
    formula <- n[,c(sp$colname)] ~ n$distance_seagrass * n$size_seagrass * n$mpa * n$depth * n$north  
  }
  print(paste("Formula",sp,":",formula))
  res <- glm(formula, family="poisson")
  print(summary(res))
  par(mfrow=c(1,2)) 
  plot(fitted(res),residuals(res),xlab="Fitted values", ylab="Residuals", abline(h=0, lty=2))
  tryCatch( 
    lines(smooth.spline(fitted(res), residuals(res))), 
    error=function(e){})
    qqnorm(res$residuals)
  qqline(res$residuals) }
```  

##GLM family=Poisson on abundance_pres

Independent variables: size_seagrass, distance_seagrass, mpa, depth, north

```{r glm_Pois_nur_loop} 

speciesrows <- c(1:3,25:28,36,39,40,42,47,52:54,58,66,84:86,88,91:100,102:104)
for (spnum in speciesrows){
  sp <- df.species_all[spnum,]
  print(sp$colname)
  n <- subset(df.sample_data, df.sample_data[c(sp$colname_pres)] > 0)
  if(nrow(n)<5){
    print("Less than 5 rows remaining in subset")
    next
  }
  if(length(unique(n$depth))==1 & length(unique(n$mpa))>1){
    formula <- n[,c(sp$colname)] ~ n$distance_seagrass * n$size_seagrass * n$mpa * n$north
  }
  else if (length(unique(n$depth))>1 & length(unique(n$mpa))==1){
    formula <- n[,c(sp$colname)] ~ n$distance_seagrass * n$size_seagrass * n$depth * n$north  
  }
  else if (length(unique(n$depth))==1 & length(unique(n$mpa))==1){
    formula <- n[,c(sp$colname)] ~ n$distance_seagrass * n$size_seagrass * n$north  
  }
  else{
    formula <- n[,c(sp$colname)] ~ n$distance_seagrass * n$size_seagrass * n$mpa * n$depth * n$north  
  }
  print(paste("Formula",sp,":",formula))
    res <- NULL
  tryCatch(
  res <- glm(formula, family="poisson"),
    error=function(e){ print("GLM kon niet uitgevoerd worden.");print(geterrmessage())}
  )
  if(is.null(res)){
    next
  }
  
  coefs <- as.data.frame(coef(summary(res)))
  coefs <- subset(coefs,coefs$`Pr(>|z|)`<=0.05)
  coefs <- data.frame(species=rep(sp,nrow(coefs)),coefficient=rownames(coefs),pvalue=coefs$`Pr(>|z|)`)
  coefs <- subset(coefs, coefs$coefficient != '(Intercept)')
  glm_tab <- rbind(glm_tab, coefs)
  if(nrow(coefs)==0){
    next
  }
  print(paste("AIC",sp,": ",summary(res)$aic))
  par(mfrow=c(1,2)) 
  plot(fitted(res),residuals(res),xlab="Fitted values", ylab="Residuals", abline(h=0, lty=2))
  title(sp)
  tryCatch( 
    lines(smooth.spline(fitted(res), residuals(res))), 
    error=function(e){}  )
  qqnorm(res$residuals)
  qqline(res$residuals) 
}
```

[1] "abudefduf_vaigiensis"
[1] "Formula Abudefduf vaigiensis : ~"                                                                     
[2] "Formula Indo-Pacific Sergeant : n[, c(sp$colname)]"                                                   
[3] "Formula TRUE : n$distance_seagrass * n$size_seagrass * n$mpa * n$depth * n$north"                     
[4] "Formula indopacificsergeant : ~"                                                                      
[5] "Formula abudefduf_vaigiensis : n[, c(sp$colname)]"                                                    
[6] "Formula abudefduf_vaigiensis_pres : n$distance_seagrass * n$size_seagrass * n$mpa * n$depth * n$north"
Error in rbind(deparse.level, ...) : 
  numbers of columns of arguments do not match

```{r glm_Pois_nur_loop_table}
knitr::kable(glm_tab, escape=TRUE)
```  

 log(size)

```{r glm_Pois_Lsize_nur_loop}

glm_tab <- data.frame(species=c(), coefficient=c(), pvalue=c())

for (sp in glm_nur_species){
  print(sp)

  res <- NULL
  tryCatch(
    res <- glm(df.sample_data[,c(sp)] ~ df.sample_data$distance_seagrass * (log(df.sample_data$size_seagrass)) * df.sample_data$mpa * df.sample_data$depth * df.sample_data$north, family="poisson"),
    error=function(e){ print("GLM kon niet uitgevoerd worden.");print(geterrmessage())}
  )
  if(is.null(res)){
    next
  }
  
  coefs <- as.data.frame(coef(summary(res)))
  coefs <- subset(coefs,coefs$`Pr(>|z|)`<=0.05)
  coefs <- data.frame(species=rep(sp,nrow(coefs)),coefficient=rownames(coefs),pvalue=coefs$`Pr(>|z|)`)
  coefs <- subset(coefs, coefs$coefficient != '(Intercept)')
  glm_tab <- rbind(glm_tab, coefs)
  if(nrow(coefs)==0){
    next
  }
  print(paste("AIC",sp,": ",summary(res)$aic))
  par(mfrow=c(1,2)) 
  plot(fitted(res),residuals(res),xlab="Fitted values", ylab="Residuals", abline(h=0, lty=2))
  title(sp)
  tryCatch( 
    lines(smooth.spline(fitted(res), residuals(res))), 
    error=function(e){}  )
  qqnorm(res$residuals)
  qqline(res$residuals) 
}
```

```{r glm_Pois_Lsize_nur_loop_table}
knitr::kable(glm_tab, escape=TRUE)
``` 

log(distance)

```{r glm_Pois_Ldist_nur_loop}

glm_tab <- data.frame(species=c(), coefficient=c(), pvalue=c())

for (sp in glm_nur_species){
  print(sp)

  res <- NULL
  tryCatch(
    res <- glm(df.sample_data[,c(sp)] ~ (log(df.sample_data$distance_seagrass)) * df.sample_data$size_seagrass * df.sample_data$mpa * df.sample_data$depth * df.sample_data$north, family="poisson"),
    error=function(e){ print("GLM kon niet uitgevoerd worden.");print(geterrmessage())}
  )
  if(is.null(res)){
    next
  }
  
  coefs <- as.data.frame(coef(summary(res)))
  coefs <- subset(coefs,coefs$`Pr(>|z|)`<=0.05)
  coefs <- data.frame(species=rep(sp,nrow(coefs)),coefficient=rownames(coefs),pvalue=coefs$`Pr(>|z|)`)
  coefs <- subset(coefs, coefs$coefficient != '(Intercept)')
  glm_tab <- rbind(glm_tab, coefs)
  if(nrow(coefs)==0){
    next
  }
  print(paste("AIC",sp,": ",summary(res)$aic))
  par(mfrow=c(1,2)) 
  plot(fitted(res),residuals(res),xlab="Fitted values", ylab="Residuals", abline(h=0, lty=2))
  title(sp)
  tryCatch( 
    lines(smooth.spline(fitted(res), residuals(res))), 
    error=function(e){}  )
  qqnorm(res$residuals)
  qqline(res$residuals) 
}
```

```{r glm_Pois_Ldist_nur_loop_table}
knitr::kable(glm_tab, escape=TRUE)
``` 

log(size) & log(distance)

```{r glm_Pois_Lsize-dist_nur_loop}

glm_tab <- data.frame(species=c(), coefficient=c(), pvalue=c())

for (sp in glm_nur_species){
  print(sp)

  res <- NULL
  tryCatch(
    res <- glm(df.sample_data[,c(sp)] ~ (log(df.sample_data$distance_seagrass)) * (log(df.sample_data$size_seagrass)) * df.sample_data$mpa * df.sample_data$depth * df.sample_data$north, family="poisson"),
    error=function(e){ print("GLM kon niet uitgevoerd worden.");print(geterrmessage())}
  )
  if(is.null(res)){
    next
  }
  
  coefs <- as.data.frame(coef(summary(res)))
  coefs <- subset(coefs,coefs$`Pr(>|z|)`<=0.05)
  coefs <- data.frame(species=rep(sp,nrow(coefs)),coefficient=rownames(coefs),pvalue=coefs$`Pr(>|z|)`)
  coefs <- subset(coefs, coefs$coefficient != '(Intercept)')
  glm_tab <- rbind(glm_tab, coefs)
  if(nrow(coefs)==0){
    next
  }
  print(paste("AIC",sp,": ",summary(res)$aic))
  par(mfrow=c(1,2)) 
  plot(fitted(res),residuals(res),xlab="Fitted values", ylab="Residuals", abline(h=0, lty=2))
  title(sp)
  tryCatch( 
    lines(smooth.spline(fitted(res), residuals(res))), 
    error=function(e){}  )
  qqnorm(res$residuals)
  qqline(res$residuals) 
}
```

```{r glm_Pois_Lsize-distance_nur_loop_table}
knitr::kable(glm_tab, escape=TRUE)
``` 

##GLM family=Gaussian on abundance with log(abundance)

log(abundance)

```{r glm_Labun_nur_loop}

glm_tab <- data.frame(species=c(), coefficient=c(), pvalue=c())

glm_nur_species <- df.species_all$colname[c(1:3,25:28,36,39,40,42,47,52:54,58,66,84:86,88,91:100,102:104)]
for (sp in glm_nur_species){
  print(sp)

  res <- NULL
  tryCatch(
    res <- glm((log(df.sample_data[,c(sp)])) ~ df.sample_data$distance_seagrass * df.sample_data$size_seagrass * df.sample_data$mpa * df.sample_data$depth * df.sample_data$north),
    error=function(e){ print("GLM kon niet uitgevoerd worden.");print(geterrmessage())}
  )
  if(is.null(res)){
    next
  }
  
  coefs <- as.data.frame(coef(summary(res)))
  coefs <- subset(coefs,coefs$`Pr(>|z|)`<=0.05)
  coefs <- data.frame(species=rep(sp,nrow(coefs)),coefficient=rownames(coefs),pvalue=coefs$`Pr(>|z|)`)
  coefs <- subset(coefs, coefs$coefficient != '(Intercept)')
  glm_tab <- rbind(glm_tab, coefs)
  if(nrow(coefs)==0){
    next
  }
  print(paste("AIC",sp,": ",summary(res)$aic))
  par(mfrow=c(1,2)) 
  plot(fitted(res),residuals(res),xlab="Fitted values", ylab="Residuals", abline(h=0, lty=2))
  title(sp)
  tryCatch( 
    lines(smooth.spline(fitted(res), residuals(res))), 
    error=function(e){}  )
  qqnorm(res$residuals)
  qqline(res$residuals) 
}
```

```{r glm_Labun_nur_loop_table}
knitr::kable(glm_tab, escape=TRUE)
```  

 log(abundance) & log(size)

```{r glm_Labun-size_nur_loop}

glm_tab <- data.frame(species=c(), coefficient=c(), pvalue=c())

for (sp in glm_nur_species){
  print(sp)

  res <- NULL
  tryCatch(
    res <- glm((log(df.sample_data[,c(sp)])) ~ df.sample_data$distance_seagrass * (log(df.sample_data$size_seagrass)) * df.sample_data$mpa * df.sample_data$depth * df.sample_data$north),
    error=function(e){ print("GLM kon niet uitgevoerd worden.");print(geterrmessage())}
  )
  if(is.null(res)){
    next
  }
  
  coefs <- as.data.frame(coef(summary(res)))
  coefs <- subset(coefs,coefs$`Pr(>|z|)`<=0.05)
  coefs <- data.frame(species=rep(sp,nrow(coefs)),coefficient=rownames(coefs),pvalue=coefs$`Pr(>|z|)`)
  coefs <- subset(coefs, coefs$coefficient != '(Intercept)')
  glm_tab <- rbind(glm_tab, coefs)
  if(nrow(coefs)==0){
    next
  }
  print(paste("AIC",sp,": ",summary(res)$aic))
  par(mfrow=c(1,2)) 
  plot(fitted(res),residuals(res),xlab="Fitted values", ylab="Residuals", abline(h=0, lty=2))
  title(sp)
  tryCatch( 
    lines(smooth.spline(fitted(res), residuals(res))), 
    error=function(e){}  )
  qqnorm(res$residuals)
  qqline(res$residuals) 
}
```

```{r glm_Labun-size_nur_loop_table}
knitr::kable(glm_tab, escape=TRUE)
``` 

log(abundance) and log(distance)

```{r glm_Labun-dist_nur_loop}

glm_tab <- data.frame(species=c(), coefficient=c(), pvalue=c())

for (sp in glm_nur_species){
  print(sp)

  res <- NULL
  tryCatch(
    res <- glm((log(df.sample_data[,c(sp)])) ~ (log(df.sample_data$distance_seagrass)) * df.sample_data$size_seagrass * df.sample_data$mpa * df.sample_data$depth * df.sample_data$north),
    error=function(e){ print("GLM kon niet uitgevoerd worden.");print(geterrmessage())}
  )
  if(is.null(res)){
    next
  }
  
  coefs <- as.data.frame(coef(summary(res)))
  coefs <- subset(coefs,coefs$`Pr(>|z|)`<=0.05)
  coefs <- data.frame(species=rep(sp,nrow(coefs)),coefficient=rownames(coefs),pvalue=coefs$`Pr(>|z|)`)
  coefs <- subset(coefs, coefs$coefficient != '(Intercept)')
  glm_tab <- rbind(glm_tab, coefs)
  if(nrow(coefs)==0){
    next
  }
  print(paste("AIC",sp,": ",summary(res)$aic))
  par(mfrow=c(1,2)) 
  plot(fitted(res),residuals(res),xlab="Fitted values", ylab="Residuals", abline(h=0, lty=2))
  title(sp)
  tryCatch( 
    lines(smooth.spline(fitted(res), residuals(res))), 
    error=function(e){}  )
  qqnorm(res$residuals)
  qqline(res$residuals) 
}
```

```{r glm_Labun-dist_nur_loop_table}
knitr::kable(glm_tab, escape=TRUE)
``` 

log(abundance), log(size) & log(distance)

```{r glm_Labun_size-dist_nur_loop}

glm_tab <- data.frame(species=c(), coefficient=c(), pvalue=c())

for (sp in glm_nur_species){
  print(sp)

  res <- NULL
  tryCatch(
    res <- glm((log(df.sample_data[,c(sp)])) ~ (log(df.sample_data$distance_seagrass)) * (log(df.sample_data$size_seagrass)) * df.sample_data$mpa * df.sample_data$depth * df.sample_data$north),
    error=function(e){ print("GLM kon niet uitgevoerd worden.");print(geterrmessage())}
  )
  if(is.null(res)){
    next
  }
  
  coefs <- as.data.frame(coef(summary(res)))
  coefs <- subset(coefs,coefs$`Pr(>|z|)`<=0.05)
  coefs <- data.frame(species=rep(sp,nrow(coefs)),coefficient=rownames(coefs),pvalue=coefs$`Pr(>|z|)`)
  coefs <- subset(coefs, coefs$coefficient != '(Intercept)')
  glm_tab <- rbind(glm_tab, coefs)
  if(nrow(coefs)==0){
    next
  }
  print(paste("AIC",sp,": ",summary(res)$aic))
  par(mfrow=c(1,2)) 
  plot(fitted(res),residuals(res),xlab="Fitted values", ylab="Residuals", abline(h=0, lty=2))
  title(sp)
  tryCatch( 
    lines(smooth.spline(fitted(res), residuals(res))), 
    error=function(e){}  )
  qqnorm(res$residuals)
  qqline(res$residuals) 
}
```

```{r glm_Labun-size-distance_nur_loop_table}
knitr::kable(glm_tab, escape=TRUE)
```