---
title: "Effect of nursery habitat availability to coral reef fish species"
author: 'Dirkje Verhoeven, Dolf Andringa, Annelies Andringa: Marine Conservation Philippines'
date: "April 9, 2016"
output:
  html_document: default
  word_document:
    fig_caption: yes
    fig_width: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readODS)
library(stringr)
library(plyr)
```
Read the spreadsheets (open document spreadsheet) and make a copy that we can use for processing the data.
```{r readfile}
df.raw <- read_ods(path='../Data_Surveys_dirkje.ods',sheet=2,col_types = c(text,numeric,numeric,numeric,numeric,numeric,numeric,numeric,numeric,date))
df.species <- as.data.frame(read_ods(path='../indicator_species.ods',sheet=1,col_types = c(text,text)))
df.species_dirkje_only <- as.data.frame(read_ods(path='../indicator_species.ods',sheet=2,col_types = c(text,text)))
df.species_nursery <- as.data.frame(read_ods(path='../indicator_species.ods',sheet=3,col_types = c(text,text)))
df.species_all <- rbind(df.species, df.species_dirkje_only)
df.processed <- df.raw

sites <- c('Antulang','Guinsuan','Malatapay','Andulay','Basak','Lutoban South','Lutoban Pier','Kookoos Nest','Dauin Poblacion')
depth <- c("5m","10m")
observer <- c('Dolf','Annelies','Dirkje','Jorien','Kallum','Emily','Melina','Rosaly','Jess','Jelle','Anna')
```

Check the summary of the data. As you can see below, all variables are text variables which they shouldn't be.
```{r summary_pre_correction}
summary(df.raw)
```

Next we'll correct the values next. First convert the DATE column to a Date value.
Then loop over the variables with integers (whole numbers) and correct those to integers.
Lastly, the variables that are actually grouping variables (factors) should be delared as such. So loop over those variables and correct them too.
```{r correct_values}


#Set the date value as a Date.
df.processed$DATE=as.Date(df.processed$DATE)

for (col in c('0-20','25-45','50-70','75-95','LOCATION','DEPTH','OBSERVER')){
  df.processed[,c(col)] <- as.integer(df.processed[,c(col)])
}
df.processed$LOCATION <- factor(df.processed$LOCATION,c(1,2,3,4,5,6,7,8,9), labels=sites)
df.processed$DEPTH <- factor(df.processed$DEPTH,c(1,2), labels=depth)
df.processed$OBSERVER <- factor(df.processed$OBSERVER,c(1,2,3,4,5,6,7,8,9,10,11), labels=observer)
df.processed$SPECIES <- factor(df.processed$SPECIES)
```
And get another summary.
```{r summary_post_correction}
summary(df.processed)
```


## Data mangling to get the species right.

```{r species_mangling}

#for the species list, make a new column with the common name where all spaces and - are removed and it is converted to lower case.
#This it is easier ti match the species names without problems with different spelling.

replace_species <- data.frame(wrong=c('Midgnight snapper',"checkered snapper","Diagonal banded sweetlip","linecheecked wrasse","Linedcheecked Wrasse","Sixtripe wrasse","Splendid soldier"), right=c('Midnight snapper','Chequered Snapper',"Diagonal banded sweetlips",'Linedcheeked Wrasse','Linedcheeked Wrasse','Sixstripe Wrasse','Splendid Soldierfish'))

df.processed$SPECIES <- mapvalues(df.processed$SPECIES, as.character(replace_species$wrong), as.character(replace_species$right))

df.species_all$common_name_mangled <- str_to_lower(str_replace_all(str_replace_all(str_replace_all(df.species_all$common_name,' ',''),'-',''),'\'',''))
df.species_all$colname <- str_to_lower(str_replace(df.species_all$scientific_name,' ','_'))

#Do the same with the imported data.
df.processed$common_name_mangled <- str_to_lower(str_replace_all(str_replace_all(str_replace_all(df.processed$SPECIES,' ',''),'-',''),'\'',''))

#Join the indicator species list to the imported data so we also have the scientific name and other info in there.
df.processed<- merge(x = df.processed, y = df.species_all, by = "common_name_mangled", all.x = TRUE)

#Get all species that don't contain a scientific name. There's probably a mis spelling with the common name.
unique(df.processed[is.na(df.processed$scientific_name),c('SPECIES')])

```


## Convert data to columns.
Create a dataframe with all unique samples. A sample is unique if the location, depth and date is unique, and then we split up the sample according to the transect number (1=0-20m, 2=25-45m, etc). We then want to add the data per species for each sample/transect combination.
```{r data_transform_prepare}

#a mapping of the transect names to the corresponding columns in the excel sheet
transects <- list('0-20','25-45','50-70','75-95')

df.sample_data <- merge(unique(df.processed[,c('LOCATION','DEPTH','DATE')]),c(1,2,3,4))
names(df.sample_data) <- c('location','depth','date','transect') #set nice column names
```
Next we create a function that gets a row from the df.sample_data dataframe (unique sample/transect) and a species name.
It selects all rows from df.processed for that sample/species and gets the number of indiduals for the correct transect. 
It sums the number of individuals because there might be multiple observers in the sample/species/transect.

```{r data_transform_function}
get_species_data <-function(row, sp){
  #select the rows from df.processed for the sample of the current row and current species
  spec_rows <- subset(df.processed,df.processed$colname == sp & df.processed$DEPTH == row['depth'] & df.processed$LOCATION == row['location'] & df.processed$DATE == row['date'])
  if(nrow(spec_rows)==0){
    #If there are no rows, set the number of fish to 0
    num = 0
  }
  else{
    #Else set the number of fish to the total for this sample/transect (sum for all observers together)
    tname <- transects[[as.integer(row['transect'])]] #translate the transect number (1,2,3,4) to the correct column name (0-20, etc)
    num = sum(spec_rows[tname])
  }
  return(num)
}
```

Loop over the species and call our function made above for each row/species, adding the species data as a column to df.sample_data.
```{r data_transform_loop}

for (sp in sort(df.species_all$colname)){
  df.sample_data[c(sp)] <- apply(df.sample_data, MARGIN=1, function(x){get_species_data(x,sp)})
}

```
## Print the data
```{r display_data, message=FALSE, echo=FALSE, eval=TRUE}
knitr::kable(df.sample_data)
```

## Save the data in a file that we can import in different R analyses.
```{r save}
save(df.sample_data,df.species_all,df.species_nursery,list=c("df.sample_data","df.species_all","df.species_nursery"), file="data.RDa")
```
